worldmap

### 2024-12-03 analysis/05_model_metrics.Rmd
Add jaccard hind cast model and figures. Move mantel test stuff.

model_data <- metrics %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "Jaccard_next",
                       "FRic", "FEve", "FDiv", "S","CWM_bodymass_value","CWM_diet_inv", 
                       "CWM_diet_scav", "CWM_diet_vfish", "CWM_diet_vunk", "CWM_forstrat_ground", 
                       "CWM_forstrat_wataroundsurf", "CWM_forstrat_watbelowsurf", "CWM_diet_fruit", 
                       "CWM_diet_planto", "CWM_diet_seed", "CWM_diet_vend", "CWM_forstrat_understory", 
                       "CWM_diet_vect", "CWM_diet_nect", "CWM_forstrat_aerial", "CWM_forstrat_canopy", 
                       "CWM_forstrat_midhigh", "CWM_age_at_maturity_max_y", "CWM_age_at_maturity_min_y", 
                       "CWM_body_size_mm", "CWM_litter_size_max_n", "CWM_litter_size_min_n", 
                       "CWM_longevity_max_y", "CWM_offspring_size_max_mm", "CWM_offspring_size_min_mm"),
         !rarefyID %in% missing_null) %>%
  group_by(metric) %>%
  mutate(study_id =  str_extract(rarefyID, "[^_]+")) %>%
  mutate(year_scaled = scale(year), 
         scale_center = attributes(year_scaled)$`scaled:center`, 
         scale = attributes(year_scaled)$`scaled:scale`) %>%
  ungroup() %>%
  #this study has many low species richness observations, so FD metrics weren't calculated, just drop it
  filter(study_id != 348) %>%
  #somehow the taxa labels didn't get updated when the timeseries for multi-taxa studies were split up, do that now
   mutate(new_taxa = str_split_fixed(rarefyID, "_", n = 3)[,3],
            taxa = case_when(
            new_taxa == "bird" ~ "Birds",
               new_taxa == "mamm" ~ "Mammals", 
               TRUE ~ as.character(taxa)
           )) %>%
  select(-new_taxa)

@@

# fit models for individual studies 
mixed_indv_mod_results <- pmap_dfr(mixed_indv_model_df %>% 
                                     filter(!metric_name %in% c("Jaccard_base", "Jaccard_next")), 
                                   get_model, data = model_data)

@@

base <- metric_model_table %>%
  filter(model_id == "base", term == "year_scaled")

>>>>

model_data <- metrics %>% 
  filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_base", "Jaccard_hind",
                       "FRic", "FEve", "FDiv", "S","CWM_bodymass_value","CWM_diet_inv", 
                       "CWM_diet_scav", "CWM_diet_vfish", "CWM_diet_vunk", "CWM_forstrat_ground", 
                       "CWM_forstrat_wataroundsurf", "CWM_forstrat_watbelowsurf", "CWM_diet_fruit", 
                       "CWM_diet_planto", "CWM_diet_seed", "CWM_diet_vend", "CWM_forstrat_understory", 
                       "CWM_diet_vect", "CWM_diet_nect", "CWM_forstrat_aerial", "CWM_forstrat_canopy", 
                       "CWM_forstrat_midhigh", "CWM_age_at_maturity_max_y", "CWM_age_at_maturity_min_y", 
                       "CWM_body_size_mm", "CWM_litter_size_max_n", "CWM_litter_size_min_n", 
                       "CWM_longevity_max_y", "CWM_offspring_size_max_mm", "CWM_offspring_size_min_mm"),
         !rarefyID %in% missing_null) %>%
  group_by(metric) %>%
  mutate(study_id =  str_extract(rarefyID, "[^_]+")) %>%
  mutate(year_scaled = scale(year), 
         scale_center = attributes(year_scaled)$`scaled:center`, 
         scale = attributes(year_scaled)$`scaled:scale`) %>%
  ungroup() %>%
  #this study has many low species richness observations, so FD metrics weren't calculated, just drop it
  filter(study_id != 348) %>%
  #somehow the taxa labels didn't get updated when the timeseries for multi-taxa studies were split up, do that now
   mutate(new_taxa = str_split_fixed(rarefyID, "_", n = 3)[,3],
            taxa = case_when(
            new_taxa == "bird" ~ "Birds",
               new_taxa == "mamm" ~ "Mammals", 
               TRUE ~ as.character(taxa)
           )) %>%
  select(-new_taxa)

@@

# fit models for individual studies 
mixed_indv_mod_results <- pmap_dfr(mixed_indv_model_df %>% 
                                     filter(!metric_name %in% c("Jaccard_base", "Jaccard_hind")), 
                                   get_model, data = model_data)

@@

base <- metric_model_table %>%
  filter(model_id == "base", term == "year_scaled",
         metric != "Jaccard_hind")

### 2024-12-03 analysis/figures/visualization.Rmd
Add jaccard hind cast model and figures. Move mantel test stuff.

plot_metric <- function(metric_name, ylabel, yval, xlabel, data, model_coef, color_list){

@@

  trend_plot <- data %>%
    left_join(pred, by = "year_scaled") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = !!sym(yval))) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey", size = 0.25) +
    geom_smooth(aes(color = climate, group = study_id), method = "glm", se = FALSE, size = 0.7) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme_classic() + 
    #theme_classic(base_size = 12) + 
    #theme(legend.position = "none") +
    theme(axis.title = element_text(size=7), legend.position = "none") +
    ylab(ylabel) +
    xlab(xlabel) +
    labs(color = "Climate") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
    scale_colour_manual(values = color_list, drop = FALSE)

@@

>>>>

plot_metric <- function(metric_name, ylabel, yval, xlabel, data, model_coef, color_list, legend){

@@

  trend_plot <- data %>%
    left_join(pred, by = "year_scaled") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = !!sym(yval))) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey", size = 0.25) +
    geom_smooth(aes(color = climate, group = study_id), method = "glm", se = FALSE, size = 0.7) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme_classic() + 
    #theme_classic(base_size = 12) + 
    #theme(legend.position = "none") +
    ylab(ylabel) +
    xlab(xlabel) +
    labs(color = "Climate") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
    scale_colour_manual(values = color_list, drop = FALSE)

  if (legend == TRUE){
    trend_plot <- trend_plot + theme(axis.title = element_text(size=7)) 
  } else{
    trend_plot <- trend_plot + theme(axis.title = element_text(size=7), legend.position = "none")
  }

@@

Get hindcast Jaccard plot for supplement
```{r}
jaccard_hind_plot <- plot_metric("Jaccard_hind", ylabel = "log(Jaccard + 1)", yval = "logvalue", xlabel = "year", data = model_data %>% mutate("log(value)" = log(value)), 
            model_coef = metric_model_table %>% filter(model_id == "base"), 
            color_list = group_colors,
            legend = TRUE)

jaccard_hind_plot + draw_grob(legend,  2/3.3, -0.3,.3/3.3, 1, scale = 0.01)

ggsave(here::here("figures", "jaccard_hind_plot.jpeg"), bg = "white", width = 6, height = 4)
```
